---
title: "Programming assigment 4"
subtitle: ""
author: "Ivan Andreu Rascon"
institute: "Rutgers"
date: "2023/3/21 (updated: `r Sys.Date()`)"
output: 
  xaringan::moon_reader:
    css: [default,hygge,rutgers,rutgers-fonts]
editor_options: 
  chunk_output_type: console
---
```{r setup, include=FALSE, warning=FALSE}
library(ds4ling)
library(tidyverse)
library(here)
library(dplyr)
library(gtsummary)
library(gt)
library(htmltools)
library(gridExtra)
library(broom)

```


```{r 4Load the dataset from inside your, echo=FALSE, warning=FALSE}

fricatives <- read.csv(here("data_raw", "fricatives.csv"))

```

```{r echo=FALSE, warning=FALSE}
tidyfricatives <- fricatives %>%
  pivot_longer(c("s_cog", "s_skewness", "sh_cog", "sh_skewness"), names_to = "class", values_to = "hz") %>%
  separate("class", into = c("phoneme", "metric"), sep = "_") %>%
  pivot_wider(names_from = "metric", values_from = "hz")

tidyfricatives %>%
write.csv(here("data_tidy","tidyfricatives.csv"))

```


```{r echo=FALSE, warning=FALSE}

tidyfricatives <- read.csv(here("data_tidy", "tidyfricatives.csv"))
summary(tidyfricatives)

```

---
# boxplot of center of gravity as a function of phoneme. 


```{r echo=FALSE, warning=FALSE}
tidyfricatives %>% 
  ggplot() +
    geom_boxplot(aes(x = phoneme, y = cog)) +
    labs(title = "center of gravity of phonemes /s/ and /ʃ/")

```

---
# skewness as a function of phoneme with stat_summary


```{r echo=FALSE, warning=FALSE}
tidyfricatives %>%
  ggplot(aes(x = phoneme, y = skewness)) +
  geom_jitter(width = 0.07, alpha = 0.5) +
  stat_summary(fun = "mean", color = "blue") +
  labs(title = "Skewness", x = "Phoneme", y = "Skewness")
```


---
8. Fit a model that examines center of gravity as a function of skewness for the [s] segments (hint: you will have to transform the data). Make a table of the model summary.

```{r echo=FALSE, warning=FALSE}

s_segments <- tidyfricatives %>%
  filter(phoneme == "s")
model <- lm(cog ~ skewness, data = s_segments)

summary(model)

fancy_table <- model %>%
  tbl_regression(exponentiate = FALSE) %>%
  as_gt()
print(fancy_table)

```
---

```{r echo=FALSE, warning=FALSE}
htmltools::tagList(fancy_table)

```

---

9. Make a scatter plot that illustrates the relationship in (8)

```{r echo=FALSE, warning=FALSE}

ggplot(data = s_segments, aes(x = skewness, y = cog)) +
  geom_point(color = "steelblue", size = 3, alpha = 0.7) +
  geom_smooth(method = "loess", se = TRUE, color = "darkred", linetype = "solid", size = 1.5) +
  labs(title = "Center of Gravity vs. Skewness for [s] Segments",
       x = "Skewness",
       y = "Center of Gravity") +
  theme_minimal() +
  theme(text = element_text(size = 14),
        plot.title = element_text(hjust = 0.5, face = "bold", size = 18),
        axis.title = element_text(face = "bold", size = 16),
        panel.grid.major = element_line(color = "grey90"),
        panel.grid.minor = element_line(color = "grey95"))

```

---
```{r model-diagnostics, echo=FALSE}
diagnosis(model)
```

---
# QQ Plot - Model diagnostics

```{r echo=FALSE}

model_diag <- augment(model)

qq_plot <- ggplot(model_diag, aes(sample = .std.resid)) +
  geom_qq() +
  geom_abline(slope = 1, intercept = 0, linetype = "dashed") +
  labs(title = "Normal Q-Q Plot",
       x = "Theoretical Quantiles",
       y = "Standardized Residuals") +
  theme_minimal()

qq_plot

```

---
**Results** 

This paper examines the relationship between center of gravity and skewness for the [s] phoneme using a linear regression model. Rresults indicated that skewness was a significant predictor of center of gravity (β = 3029.1, t(18) = 11.69, p < .001), accounting for 87.71% of the variance (R-squared) in center of gravity. 
The intercept (β = 8027.2, t(18) = 37.80, p < .001) represents the estimated center of gravity when skewness is equal to zero. The t-value associated with the intercept indicates that the estimated intercept is different from zero at 0.001 significance.
The residual standard error was 320.1, and indicates the average distance of the observed data points from the fitted regression line. The R-squared value of 0.8836 indicates that 88.36% of the variance in center of gravity can be explained by the model, while the adjusted R-squared value of 0.8771 adjusts for the number of predictors and degrees of freedom.

Finally, the Q-Q plot showed that the residuals were normally distributed. 

---

```{r echo=FALSE}

assumptions <- read.csv(here("data_raw", "assumptions.csv"))

```

**This is awesome**

And would not fit a linear model because:

1. The relationship between x and y is not linear.
2. Values are not normally distributed.
3. Heteroscedasticity: The variance of y is not constant across the range of x values.

```{r echo=FALSE}

assumptions %>% 
  ggplot() +
    geom_point(aes(x = x, y = y)) +
    labs(title = "Assumptions")



```

